The Real-time fMRI Neurofeedback Based Stratification of Default Network Regulation Neuroimaging Data Repository.

### 2016-10-08 reviewer_response_letter.tex

\documentclass{article}

\usepackage{lipsum}
% \usepackage{layout}

\newcommand{\ts}{\textsuperscript}
%\usepackage{fontspec}
%\setmainfont{Calibri}
\providecommand\CMItoname{}
\providecommand\CMItoaddress{}
\providecommand\CMItoopeningname{Editor}
\providecommand\CMIreline{Response to reviews for Ms. No.: NIMG-16-2197}
\input{NKIletterdefs}

\usepackage[normalem]{ulem}
\definecolor{responseblue}{HTML}{0000A0}
\newcommand{\RESPONSE}[1]{\textcolor{responseblue}{#1}}

@@

\begin{document}
\sffamily

We would like to thank the reviewers for their indepth and thoughtful critique of our manuscript titled ``Real-time fMRI Neurofeedback Based Stratification of Default Network Regulation Neuroimaging Data Repository''. We have updated the manuscript to address the criticisms and present a point by point response in \RESPONSE{blue text} below.

@@

\RESPONSE{We apologize for the confusion, but the study described herein is not the NKI-RS, but a separate experiment, with separate funding and aims. We have clarified this in the introduction: ``The Default Network Regulation Neuroimaging Repository contains data from a suite of fMRI experiments aimed at better understanding individual variation in DMN activity and modulation.  Although it is a separate project, it has been harmonized with, and is distributed alongside, the Enhanced Nathan Kline Institute-Rockland Sample(NKI-RS) [19], which aims to capture deep and broad phenotyping of a large community-ascertained sample.''}

Three block-design fMRI tasks were administered in addition to resting state scans: a Moral Dilemma task (previously shown to "activate" the DMN), an interference task, designed to recruit cognitive control networks, and a neuro-feedback task, in which participants were instructed, by visual cue, to either "focus" their thoughts or let their minds wander. The results are presented in Figs. 7 - 8. Thus, Fig. 7 demonstrates that the Moral Dilemma task activated the DMN while suppressing the anti-DMN; a generally inverse response was generated by the interference task. Regarding the resting state data, the topography of the DMN was extracted using ICA-based dual regression. Fig. 8 shows that the BOLD fMRI signal summed over the DMN was modulated as expected by the alternating mind wandering/focus task. This effect was decidedly more pronounced in data excluding participants who self-reported sleeping during the scan. Fig. 6 reports data quality measures, e.g., standardized
DVARS compiled over the various task conditions. Results correlating (across subjects) the fMRI data against phenotypic measures, i.e., the originally stated objective of the NKI-RS initiative, are conspicuously absent.

\RESPONSE{This manuscript is a data descriptor that describes a data-sharing repository. In accordance with NeuroImage's advice for publishing such manuscripts, and other journal's formats for similar publications, it includes a technical validation of the data in the repository, and explicitly excludes any new analyses or novel scientific findings. We have clarified this point in the manuscript by directly referring to it as a data descriptor in the Abstract and Introduction.}

Specific critiques of this work might include, for example, that sleep was assessed by self-report rather than fMRI-based measures (Tagliazucchi et al., 2012). That more participants slept than originally anticipated is discussed as an unexpected result. However, this is a predictable feature of studies involving community-based participants (Tagliazucchi \& Laufs, 2014).

\RESPONSE{We agree with the reviewer that sleep is a problem for large-scale fMRI experiments, which is exactly why we included a self-report measure to evaluate this phenomenon. Whether the fMRI-based mentioned prediction methods provides more accurate data than self report remains to be determined. Additionally, we include GSR, respiration, and pulse information that may be useful for objectively measuring sleep. We have added a paragraph to the \emph{Usage Notes} discussing sleep in more detail.}

\RESPONSE{``The impact of sleep on intrinsic brain activity, and the preponderance of sleep that occurs during resting state fMRI acquisition, has been highlighted in the literature  [78, 79]. Indeed, a large proportion of the participants in this  study  reported  falling  asleep during either the resting state or neurofeedback scans. Whether or not this  data should be excluded is up to the researcher and depends on the analysis being performed. Users of this resource might also consider whether they trust the self reports, or whether they should try to decode an objective measure of sleep from the data [79]. Additionally, researchers could potentially use the respiratory, heart rate, or galvanic skin response recordings provided in the repository to monitor wakefulness.''}

The principal issue is that, although the necessary data have been acquired, the authors have abrogated their original commitment to studying the functional neuroimaging correlates of psychological measures. Thus, this paper is pitched as announcing the existence of a data repository rather than as presenting scientific results. Whether this type of paper merits publication is uncertain.

@@

\RESPONSE{The reviewer is correct that the paper is a data descriptor and not meant to carry out any in-depth analyses. The appropriateness of data descriptors for publication in NeuroImage is supported by their recent special issues on data sharing as well as communications from the editor-in-chief.}

However, after 4 papers of promise, it would seem, to this reviewer at least, that the time to present some results has arrived. Surely, the grant proposal funding this work mentions potential associations between
personality traits and measurable fMRI features (resting state or task-based). Why not put some of these hypotheses to the test?

\RESPONSE{This comment seems to be related to the same lack of clarity about the relationship between the described study and the NKI-RS that we addressed above. When it comes to the impact of the greater NKI-RS effort, we have a much more optimistic view. Data shared through this initiative have been used by 122 publications, many of which report phenotypic correlations (\url{http://fcon 1000.projects.nitrc.org/indi/enhanced/pubs.html}). We draw particular attention to our recent publication in Biological Psychiatry (Van Dam et al., in press), which is directly relevant to associations between personality traits and R-fMRI features. Of note, this paper made use of multivariate distance matrix regression - a technique for which the NKI-RS datasets were previously used to carry out the first application in functional connectomics (i.e., detection of brain-behavior relationships for IQ) (Shehzad et al., 2014; Neuroimage).}

@@

\RESPONSE{The reviewer is also correct that the grant funding the DMN neurofeedback experiment does refer to brain-behavior relationships, but importantly, it also requires the data to be shared as it is collected. The goal of this data descriptor is to provide users with an in-depth description of the data, how it was collected, and a technical validation. }

Possibly, the authors have already carried out such analyses without finding statistically significant associations. If so, that does not matter. Negative findings are of interest, provided that they are accompanied by an appropriate power analysis.

\RESPONSE{We agree with the reviewer about the importance of publishing negative results, and promise to do so when they occur. Data collection for this project is ongoing and we are waiting for larger samples and replication before publishing brain-behavior correlations. But the results of these analyses are beyond the scope of a data descriptor.}

@@

Relating functional connectivity measures to cognitive performance measures, e.g., as in (Finn et al., 2015), is fraught with interpretive difficulties (Siegel et al., 2016). Possibly, the authors are aware of these developments and, hence, are understandably cautious. However, these difficulties make it even more imperative that experts (i.e., the authors) rather than amateurs conduct the required analyses, the latter option being what is implied by the notion of a data repository.

\RESPONSE{We disagree wholeheartedly with the notion that only we that collected the data are qualified to analyze them and interpret the findings. Indeed, the open access of the data through repositories such as the one we describe here make it possible for others to re-analyze the data to validate our results and interpretations (for example, Siegel et al., 2016 and Tagliazucchi et al., 2014). We believe that data sharing addresses the concerns raised by the reviewer more effectively than relegating analyses only to the anointed and anonymous peer review.}

### 2016-10-08 mcdonald_nfb_datasharing_abstract.tex
more edits to address reviewer comments and response to reviews

This paper describes a repository of openly shared data from an experiment to assess inter-individual differences in default mode network (DMN) activity. This repository includes cross-sectional functional magnetic resonance imaging (fMRI) data from the Multi Source Interference Task, to assess DMN deactivation, the Moral Dilemma Task, to assess DMN activation, a resting state fMRI scan, and a DMN neurofeedback paradigm, to assess DMN modulation, along with accompanying behavioral and cognitive measures. We report technical validation from n=125 participants of the final targeted sample of 180 participants. Each session includes acquisition of one whole-brain anatomical scan and whole-brain echo-planar imaging (EPI) scans, acquired during the aforementioned tasks and resting state. The data includes several self-report measures related to perseverative thinking, emotion regulation, and imaginative processes, along with a behavioral measure of rapid visual information processing.

>>>>

This data descriptor describes a repository of openly shared data from an experiment to assess inter-individual differences in default mode network (DMN) activity. This repository includes cross-sectional functional magnetic resonance imaging (fMRI) data from the Multi Source Interference Task, to assess DMN deactivation, the Moral Dilemma Task, to assess DMN activation, a resting state fMRI scan, and a DMN neurofeedback paradigm, to assess DMN modulation, along with accompanying behavioral and cognitive measures. We report technical validation from n=125 participants of the final targeted sample of 180 participants. Each session includes acquisition of one whole-brain anatomical scan and whole-brain echo-planar imaging (EPI) scans, acquired during the aforementioned tasks and resting state. The data includes several self-report measures related to perseverative thinking, emotion regulation, and imaginative processes, along with a behavioral measure of rapid visual information processing.

### 2016-10-08 mcdonald_nfb_datasharing_body.tex
more edits to address reviewer comments and response to reviews

Herein, we provide an overview of planned data collection, methods used, summaries of data collected and available to date, and validation analyses. New data will be released on a regular basis and will be available at the Collaborative Informatics and Neuroimaging Suite (COINS) Data Exchange (\url{http://coins.mrn.org/dx}) \cite{Scott2011,Wood2014}, as well as the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC; \url{http://www.nitrc.org/}) and in the Amazon Web Services (\url{https://aws.amazon.com/s3/}).

@@

For the MSIT, fMRI data should be analyzed as 42s blocks with the following onset times after dropping the first four TRs: congruent blocks - 22, 106, 190, 274; incongruent blocks –64,148, 232, 316. The Moral Dilemma fMRI data should be analyzed as 30s blocks with the following onset times after dropping the first four TRs: control blocks – 12, 72, 132, 192; dilemma blocks – 42, 102, 162, 222. Note that up until April 2014 (first 11 participants), no pre-experiment fixation period was implemented for the moral dilemma task. For those participants, all of the aforementioned stimulus onset times should be altered by subtracting 20s. For analysis of the NFB fMRI data, the first stimulus onset (and duration) times, in seconds, are as follows: 34(30), 162(60), 260(90), 418(60), 576(30), 674(90). The second stimulus onset (and duration) times, in seconds, are as follows: 68(90), 226(30), 354(60), 482(90), 610(60), 768(30).

@@

@@

One unexpected result of our technical validation is a high amount of data loss due to poor participant performance and compliance. Excluding all of the participants who are considered outliers on at least one of the tasks or who fell asleep during the resting state or neurofeedback scans, would remove 73 of the 125 participants (see Fig.~\ref{fig:outliers}). Forty-six individuals could be excluded for sleep, which is not surprising given recent reports of the high incidence of sleep during resting state scans \cite{Tagliazucchi2014}. Interestingly, 26 of these participants were also outliers in at least one other task. This indicates that either they had trouble staying awake during the other tasks as well, or were non-compliant overall. Due to a paucity of information on participant compliance during fMRI experiments, it is hard to tell whether what we are seeing is common for our population, or whether there is a specific problem with our experiment protocol. We do believe that compliance would be higher if we were to utilize a younger healthy population, or participants who have been scanned multiple times, as is commonly done in cognitive neuroscience. These problems with sleep and poor performance illustrate the need to debrief participants or otherwise monitor their wakefulness during the scan. Additionally, it is important to check the data as it is acquired so that a study protocol can be adapted to reduce data loss.

>>>>

In this data descriptor, we provide an overview of planned data collection, methods used, summaries of data collected and available to date, and validation analyses. New data will be released on a regular basis and will be available at the Collaborative Informatics and Neuroimaging Suite (COINS) Data Exchange (\url{http://coins.mrn.org/dx}) \cite{Scott2011,Wood2014}, as well as the Neuroimaging Informatics Tools and Resources Clearinghouse (NITRC; \url{http://www.nitrc.org/}) and in the Amazon Web Services (\url{https://aws.amazon.com/s3/}).

@@

For the MSIT, fMRI data should be analyzed as 42s blocks with the following onset times after dropping the first four TRs: congruent blocks - 22, 106, 190, 274; incongruent blocks –64,148, 232, 316. The Moral Dilemma fMRI data should be analyzed as 30s blocks with the following onset times after dropping the first four TRs: control blocks – 12, 72, 132, 192; dilemma blocks – 42, 102, 162, 222. Note that up until April 2014 (first 11 participants), no pre-experiment fixation period was implemented for the moral dilemma task. For those participants, all of the aforementioned stimulus onset times should be altered by subtracting 20s. For analysis of the NFB fMRI data, the first stimulus onset (and duration) times, in seconds, are as follows: 34(30), 162(60), 260(90), 418(60), 576(30), 674(90). The second stimulus onset (and duration) times, in seconds, are as follows: 68(90), 226(30), 354(60), 482(90), 610(60), 768(30). This timing information is also available in \textttt{events.tsv} files provided in the repository alongside the task fMRI data.

@@

The  impact of sleep on intrinsic brain activity, and the preponderance of sleep that occurs during resting state fMRI acquisition, has been highlighted in the literature \cite{Duyn2011,Tagliazucchi2014}. Indeed, a large proportion of the participants in this study reported falling asleep during either the resting state or neurofeedback scans. Whether or not this data should be excluded is up to the researcher and depends on the analysis being performed. Users of this resource might also consider whether they trust the self reports, or whether they should try to decode an objective measure of sleep from the data \cite{Tagliazucchi2014}. Additionally, researchers could potentially use the respiratory, heart rate, or galvanic skin response recordings provided in the repository to monitor wakefulness.

@@

One unexpected result of our technical validation is a high amount of data loss due to poor participant performance and compliance. Excluding all of the participants who are considered outliers on at least one of the tasks or who fell asleep during the resting state or neurofeedback scans, would remove 73 of the 125 participants (see Fig.~\ref{fig:outliers}). Forty-six individuals could be excluded for sleep, which is not surprising given recent reports of the high incidence of sleep during resting state scans \cite{Tagliazucchi2014}. Interestingly, 26 of these participants were also outliers in at least one other task. This indicates that either they had trouble staying awake during the other tasks as well, or were non-compliant overall. Due to a paucity of information on participant compliance during fMRI experiments, it is hard to tell whether what we are seeing is common for our population, or whether there is a specific problem with our experiment protocol. We do believe that compliance would be higher if we were to utilize a younger healthy population, or participants who have been scanned multiple times, as is commonly done in cognitive neuroscience. These problems with sleep and poor performance illustrate the need to debrief participants, monitor their wakefulness during the scan, or try decode sleep from the fMRI data \cite{Tagliazucchi2014}. Additionally, it is important to check the data as it is acquired so that a study protocol can be adapted to reduce data loss.
